---
title: "Homework 2"
author: "Tony Colucci/Naomi Kaduwela"
date: "January 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(fpc)
library(mclust)

library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
```

# 2
```{r}
rm(list = ls())
#read in data
nuoqiData <- read.csv("nuoqi.csv", stringsAsFactors = FALSE)
#display structure
str(nuoqiData)


#code to create k means summary

summary.kmeans = function(nuoqiKmeans) {
p = ncol(nuoqiKmeans$centers)
k = nrow(nuoqiKmeans$centers)
n = sum(nuoqiKmeans$size)
sse = sum(nuoqiKmeans$withinss)
xbar = t(nuoqiKmeans$centers)%*%nuoqiKmeans$size/n
ssb = sum(nuoqiKmeans$size*(nuoqiKmeans$centers - rep(1,k) %*% t(xbar))^2) 
print(data.frame(
n=c(nuoqiKmeans$size, n),
Pct=(round(c(nuoqiKmeans$size, n)/n,2)),
round(rbind(nuoqiKmeans$centers, t(xbar)), 2)
#RMSE = round(sqrt(c(nuoqiKmeans$withinss/(p*(nuoqiKmeans$size-1)), rss/(p*(n-k)))), 4)
))
cat("SSE = ", sse, "; SSB = ", ssb, "\n")
cat("R-Squared = ", ssb/(ssb+sse), "\n")
cat("Pseudo F = ", (ssb/(k-1))/(sse/(n-k)), "\n\n");
invisible(list(sse=sse, ssb=ssb, Rsqr=ssb/(ssb+sse), F=(ssb/(k-1))/(sse/(n-k))))
}


```

## a)
```{r}
set.seed(12345)
num_clusters = 5
nuoqiKmeans <- kmeans(nuoqiData[,1:5], centers = num_clusters, nstart = 100, iter.max = 1000)

nuoqiKmeans$centers
nuoqiKmeans$size
nuoqiKmeans$withinss
nuoqiKmeans$betweenss

summary(nuoqiKmeans)

```
#Model 1:
SSE =  813.4389 ; 
SSB =  1156.095 
R-Squared =  0.5869891 
Pseudo F =  351.4025 

```{r}
#violin plot 
plot.kmeans = function(fit,boxplot=F) {
  require(lattice)
  p = ncol(fit$centers) 
  k = nrow(fit$centers) 
  plotdat = data.frame(
    mu=as.vector(fit$centers),
    clus=factor(rep(1:k, p)),
    var=factor( 0:(p*k-1) %/% k, labels=colnames(fit$centers))
  )
  print(dotplot(var~mu|clus, data=plotdat,
                panel=function(...){ 
                  panel.dotplot(...) 
                  panel.abline(v=0, lwd=.1)
                },
                layout=c(k,1), xlab="Cluster Mean"
  )) 
  invisible(plotdat) }

plot(nuoqiKmeans)
```
Clusters 2, 4 and 5 contain people who gave approximately the same answers across all 5 questions, with cluster 2 having those people who responded lower, cluster 5 having people who responded higher and cluster 4 in the middle. Cluster 1 contains people who place high value on selfexpress.

#Cluster 1 = self express
#Cluster 2 = fashion enthusiast
#Cluster 3 = evenly spread except for cross - functional
#Cluster 4 = evenly spread - maybe cross
#Cluster 5 = evenly spaced - impressive

Within cluster sum of squares by cluster:
[1] 158.9550 192.8771 144.1307 207.2093 110.2668
 (between_SS / total_SS =  58.7 %)
 
 
```{r}
#plot 2D visual of cluster
fviz_cluster(nuoqiKmeans, data = nuoqiData)
```


#Here we see a lot of overlap in clusters


```{r}
#see how euclidean distance is across cluster means 
distance1 <- get_dist(nuoqiKmeans$centers)
fviz_dist(distance1, gradient = list(low = "green", mid = "white", high = "red"))


```



## b)
```{r}
library(dplyr)
#ipsatization 
#standardize vriables by subtracing xbar (avg response for given respondent to all 5 point scales) by their variable value
nuoqiDataIps <- mutate(nuoqiData,
                    xbar = (fashethus + cross + functional + selfexpress + impress)/5,
                    fashethusIps = fashethus - xbar,
                    crossIps = cross - xbar,
                    functionalIps = functional - xbar,
                    selfexpressIps = selfexpress - xbar,
                    impressIps = impress - xbar)
```

## c)
```{r}
set.seed(12345)
nuoqiKmeansIps <- kmeans(nuoqiDataIps[,10:14], centers = num_clusters, nstart = 100, iter.max = 1000)

nuoqiKmeansIps$centers
nuoqiKmeansIps$size
nuoqiKmeansIps$withinss
nuoqiKmeansIps$betweenss
nuoqiKmeans$tot.withinss


plot(nuoqiKmeansIps)

```
#Still hard to tell clear segments between the 5 categoris...
#Now All clusters are high on self express except cluster 4 and two is slightly less as well


#Cluster 1 = self express
#Cluster 2 = fashion enthusiast - self expressive 
#Cluster 3 = evenly spread except for cross - self express/fashion enthuasiast 
#Cluster 4 = evenly spread - maybe cross --> now might be functional
#Cluster 5 = evenly spaced - impressive --> highest in self express

```{r}
summary(nuoqiKmeansIps)

distance <- get_dist(nuoqiKmeansIps$centers)
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))

fviz_cluster(nuoqiKmeansIps, data = nuoqiDataIps)
```

#We see in model two the clusters overlap more, SSE is lower as is the SSB in model 2
#Model 1 has higher R^2 and F score

#Model 2:
SSE =  442.2794 ; 
SSB =  368.0022 
R-Squared =  0.4541658 
Pseudo F =  205.7264 

#Model 1:
SSE =  813.4389 ; 
SSB =  1156.095 
R-Squared =  0.5869891 
Pseudo F =  351.4025 


## d)
```{r}
for (n_clust in 2:6) {
  set.seed(12345)
  nuoqiKmeans <- kmeans(nuoqiData[,1:5], centers = n_clust, nstart = 100, iter.max = 1000)
  print(paste(n_clust, " cluster solution for untransformed data"))
  print(nuoqiKmeans$centers)
  print(nuoqiKmeans$size)
  print(nuoqiKmeans$withinss)
  print(nuoqiKmeans$betweenss)
}

for (n_clust in 2:6) {
  set.seed(12345)
  nuoqiKmeans <- kmeans(nuoqiDataIps[,10:14], centers = n_clust, nstart = 100, iter.max = 1000)
  print(paste(n_clust, " cluster solution for ipsatized data"))
  print(nuoqiKmeans$centers)
  print(nuoqiKmeans$size)
  print(nuoqiKmeans$withinss)
  print(nuoqiKmeans$betweenss)
}
```

```{r}
#Elbow plot
set.seed(123)
View(nuoqiData)
fviz_nbclust(nuoqiData[,1:5], kmeans, method = "wss") #total within-cluster sum of square 
fviz_nbclust(nuoqiDataIps[,10:14], kmeans, method = "wss") #total within-cluster sum of square 


```
nuoqiData
#The optimal number of clusters takes the biggest dip after 1 but around 3-4 clusters look like there's still sig decrease
#Flattening off after 5

nuoqiDataIps
#Around 3 or 5 clusters, could try 8-9 also before it levels off

```{r}

#Added visualization to compare across clusters with different sizes 
k2 <- kmeans(nuoqiData[,1:5], centers = 2, nstart = 25, iter.max = 1000)
k3 <- kmeans(nuoqiData[,1:5], centers = 3, nstart = 25, iter.max = 1000)
k4 <- kmeans(nuoqiData[,1:5], centers = 4, nstart = 25, iter.max = 1000)
k5 <- kmeans(nuoqiData[,1:5], centers = 5, nstart = 25, iter.max = 1000)
k6 <- kmeans(nuoqiData[,1:5], centers = 6, nstart = 25, iter.max = 1000)

# plots to compare
p2 <- fviz_cluster(k2, geom = "point", data = nuoqiData[,1:5]) + ggtitle("k = 2")
p3 <- fviz_cluster(k3, geom = "point", data = nuoqiData[,1:5]) + ggtitle("k = 3")
p4 <- fviz_cluster(k4, geom = "point",  data = nuoqiData[,1:5]) + ggtitle("k = 4")
p5 <- fviz_cluster(k5, geom = "point",  data = nuoqiData[,1:5]) + ggtitle("k = 5")
p6 <- fviz_cluster(k6, geom = "point",  data = nuoqiData[,1:5]) + ggtitle("k = 6")

library(gridExtra)
grid.arrange(p2, p3, p4, p5, p6, nrow = 2)



#Added visualization to compare across clusters with different sizes - ipsatized
ik2 <- kmeans(nuoqiDataIps[,10:14], centers = 2, nstart = 25, iter.max = 1000)
ik3 <- kmeans(nuoqiDataIps[,10:14], centers = 3, nstart = 25, iter.max = 1000)
ik4 <- kmeans(nuoqiDataIps[,10:14], centers = 4, nstart = 25, iter.max = 1000)
ik5 <- kmeans(nuoqiDataIps[,10:14], centers = 5, nstart = 25, iter.max = 1000)
ik6 <- kmeans(nuoqiDataIps[,10:14], centers = 6, nstart = 25, iter.max = 1000)

# plots to compare
ip2 <- fviz_cluster(ik2, geom = "point", data = nuoqiDataIps[,10:14]) + ggtitle("k = 2i")
ip3 <- fviz_cluster(ik3, geom = "point", data = nuoqiDataIps[,10:14]) + ggtitle("k = 3i")
ip4 <- fviz_cluster(ik4, geom = "point",  data = nuoqiDataIps[,10:14]) + ggtitle("k = 4i")
ip5 <- fviz_cluster(ik5, geom = "point",  data = nuoqiDataIps[,10:14]) + ggtitle("k = 5i")
ip6 <- fviz_cluster(ik6, geom = "point",  data = nuoqiDataIps[,10:14]) + ggtitle("k = 6i")

grid.arrange(ip2, ip3, ip4, ip5, ip6, nrow = 2)

```

#The first version is better because the clusters are more sepearated; K=3 cluster looks the best
#The ipsatized version has a lot of overlap between clusters which is not good for k means because it will result in misclassificaion. 


## e)
```{r}
set.seed(12345)
nuoqiGaussian <- Mclust(nuoqiData[,1:5], G=5)
summary(nuoqiGaussian)
plot(nuoqiGaussian)

set.seed(12345)
nuoqiGaussianIps <- Mclust(nuoqiDataIps[,10:14], G=5)
summary(nuoqiGaussianIps)
plot(nuoqiGaussianIps)
```


